---
title: "Assignment 4 Clustering"
author: "YASHWANTH REDDY SADALA"
date: "2023-11-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem Statement:
An equities analyst is studying the pharmaceutical industry and would like your help in exploring and understanding the financial data collected by her firm. Her main objective is to understand the structure of the pharmaceutical industry using some basic financial measures. Financial data gathered on 21 firms in the pharmaceutical industry are available in the file Pharmaceuticals.csv. For each firm, the following variables are recorded:  

1.Market capitalization (in billions of dollars)
2.Beta
3.Price/earnings ratio
4.Return on equity
5.Return on assets
6.Asset turnover
7.Leverage
8.Estimated revenue growth
9.Net profit margin
10.Median recommendation (across major brokerages)
11.Location of firm’s headquarters
12.Stock exchange on which the firm is listed

Use cluster analysis to explore and analyze the given dataset as follows: 

1.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 
2.Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
3.Provide an appropriate name for each cluster using any or all of the variables in the dataset.

#Running the necessary libraries
```{r}
library(tidyverse)
library(factoextra)
library(dplyr)
library(ggplot2)
library(cluster)
```

#Importing and reading the CSV file
```{r}
Pharmaceuticals <- read.csv("C:/Users/yashw/FML/Pharmaceuticals.csv")
Pharma <- na.omit(Pharmaceuticals) 
head(Pharmaceuticals)
dim(Pharmaceuticals)
t(t(names(Pharmaceuticals)))
```

#1. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.

```{r}
#using only the quantitative variables(1-9) to cluster the 21 firms
row.names(Pharma)<- Pharma[,1]
Pharma1<- Pharma[,3:11]

#Considering only numerical values 3-11 columns.
head(Pharma1)
```

```{r}
#Normalizing data
Pharma2<-scale(Pharma1)
head(Pharma2)
```
#Additionally, to calculate how many clusters to need for the Elbow Method cluster analysis
```{r}
fviz_nbclust(Pharma2, kmeans, method = "wss")
```
#Looking at the Elbow method's above graph, we can see that it's unclear which of the values k=2, 3, 4, or 5 to select. The silhouette method for counting the number of clusters
```{r}
fviz_nbclust(Pharma2, kmeans, method = "silhouette")
```
#Applying K-means clustering 
```{r}
set.seed(150)
clus5<- kmeans(Pharma2,centers=5,nstart = 25)
```

#Visualizing the output
```{r}
clus5$centers
```
#Visualizing the clusters
```{r}
fviz_cluster(clus5,data = Pharma2) 
```

```{r}
clus5
```

```{r}
distance<- dist(Pharma2, method = "euclidean")
fviz_dist(distance)
```

#There are five clusters, as can be seen, and the center is established after 25 restarts, as calculated using the k-means algorithm. K - Cluster Analysis of Means Fit five clusters to the data.
```{r}
fit<-kmeans(Pharma2,5)
```

#Determining each cluster's mean value for every quantitative variable
```{r}
aggregate(Pharma2,by=list(fit$cluster),FUN=mean)

```

#To see the layout of clusters
```{r}
clusplot(Pharma2,fit$cluster,color =
           TRUE,shade = TRUE,labels = 2,lines = 0)

```

#2. Interpret the clusters in light of the numerical variables that were utilised to create them.
```{r}
Pharmacluster <- Pharma[,c(12,13,14)]%>% mutate(clusters = clus5$cluster)%>% arrange(clusters, ascending = TRUE)
Pharmacluster
```

#Cluster 1 - AHM, WYE,BMY,AZN,ABV,SGP,LLY,NVS ( low Market_Cap,low Beta,low PE_Ratio,high Leverage,high Rev_Growth.)
#Cluster 2 - GSK,JNJ,MRK and PFE (High Market_Cap,ROE, ROA,Asset_Turnover Ratio and low Beta/PE Ratio)
#Cluster 3 - AGN,PHA (low Asset_Turnover, High PE Ratio)
#Cluster 4 - ELN,MRX,WPI,AVE (low PE_Ratio,high ROE,low ROA,low Net_Profit_Margin, high Rev_Growth)
#Cluster 5 - BAY,CHTT,IVX (low Rev_Growth,high Beta and levearge,low Net_Profit_Margin)

#Do the clusters exhibit any patterns in relation to the numerical variables (10–12)? (those not utilized in cluster formation)

```{r}
Pharma <- Pharmaceuticals[12:14] %>% mutate(Clusters=clus5$cluster)
ggplot(Pharma, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters',y ='Frequence')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Location))+
  geom_bar(position = 'dodge')+labs(x ='Clusters',y = 'Frequence')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+
  labs(x ='Clusters',y = 'Frequence')
```

#As can be seen from the graph above, cluster 1 has a moderate level of leverage and little profit. The graph establishes a moderate purchase and hold level.
#Because cluster 2 has an equal amount of moderate purchase and sell, there may be room for growth in these companies given their significant revenue growth.
#Cluster 3 has a high hold rate and a sizable profit margin, which will entice investors to purchase more of this cluster.
#Cluster 4 exhibits both large profit margins and a significant degree of market capitalization. It reflects the company's great potential and has a degree of purchase and hold equal to it.
#The graph in cluster 5 illustrates how high debt causes high leverage.


#3. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

#Cluster 1: Perfect Asset
#Cluster 2: Investment over the long run
#Cluster 3: Dangerous Risk 
#Cluster 4: potential Growth 
#Cluster 5: Investment over the short term

